---
title: "Problem 2.2: The Legacy-Adaptive-Legacy (LAL) diffusion model"
---
```{r setup, include=FALSE}

# Load dependencies.
source("../socmod.R")
```


## Analytical calculations with LAL (1 point)

We can do some relatively simple algebra with the recursive LAL model to 
predict whether 

### Diffusion rate at introduction when $t,~A_t \approx 0$ (0.5 points)


### Equilibrium $A_t$, $\hat A$, as $t\to\infty$ (0.5 points)

First, follow _MSB_ pp. 95-97 to show that the _prevalence_ of the behavior $A$
(i.e., the proportion of the population doing $A$)
at equilibrium is 
$$
\hat a = \frac{\hat A}{N} = 
$$

Use the mathematics in this and the previous problem set as models for writing
out in your own words the steps from . Make sure to explain why at equilibrium
$\Delta A = 0$ and $A_{t+1} = A_t = \hat A$. Why isn't $A = 1$ a stable
equilibrium? Explain in terms of  Use terminology that matches
behavioral adoption, instead of epidemiological terminology. 
Substitute the adoption rate, $\alpha$, for the
transmissibility, $\tau$, and use the drop rate $\delta$ in place of the 
recovery rate $\gamma$. 


## Agent-based LAL analysis (1 point)


### Plot several time series for different 

```{r}
lal_model_step <- function(model) {
  
}
lal_model <- AgentBasedModel$new()
```


### Heatmap of equilibrium adoption $\hat A$ over $\alpha$ and $\delta$


From our derivation above, we ahve 
