---
title: "Problem 2.2: The Legacy-Adaptive-Legacy (LAL) diffusion model"
---
```{r setup, include=FALSE}

# Load dependencies.
source("../socmod.R")
```


## Analytical calculations with LAL (1 point)

We can do some relatively simple algebra with the recursive LAL model to 
predict the equilibrium value of adoption.


### Derive the equilibrium adoption level, $\hat A$ (0.5 points)

First, follow _MSB_ Section 4.4, pp. 95-97, to show that the _prevalence_ 
of the behavior $A$ (i.e., the proportion of the population doing $A$) at 
equilibrium is 
$$
\frac{\hat A}{N} = 1 - \frac{\delta}{\alpha}.
$$

Use the source document mathematics in this, the previous problem set, and
see also the [source document for the Course
Notes](https://mt.digital/teaching/CSS4S/notes/) as models for writing
mathematics using \LaTeX commands. 


### Explaining stable and unstable equilibria (0.5 points)

Next, explain why equilibrium means that $\Delta A = 0$, again using _MSB_
Section 4.4. Use terminology that matches
the adoption of adaptive behaviors instead of epidemiological terminology. 
Substitute the adoption rate, $\alpha$, for the
transmissibility, $\tau$, and use the drop rate $\delta$ in place of the 
recovery rate $\gamma$. 

- Make sure to specifically address why $A = \hat A$ is a _stable_ equilibrium,
  but $A = 0$ is not, even though $\Delta A = 0$ when $A = 0$. 
- Support your explanation using the two plots from Figure 4.9 using the `hline`
  function in `ggplot2` to draw the horizontal line at 0.0 in your plot of
  $\Delta A$ over $A$ (left plot in Figure 4.9 of *MSB*) and at $A = \hat A =
  0.75$ (right plot).


## Agent-based LAL analysis (1 point)

Now 

### Plot several time series for different 

```{r}
lal_model_step <- function(model) {
   
}
lal_model <- AgentBasedModel$new()
```

### 
